---
title: "What are the Most Cost-Effective Ways to Fight Climate Change?"
---

```{r libraries,echo=FALSE,results=FALSE,warning=FALSE,error=FALSE,message=FALSE}
library(gsheet)
library(metafor)
#library(RMariaDB)
library(RMySQL)
library(knitr)
library(ggplot2)
library(dplyr)
```

```{r source,echo=FALSE,results=FALSE,warning=FALSE,error=FALSE,message=FALSE}
source('source.R')
read_chunk('SKY_core_functions.R')
SKY.db <- dbConnect(MySQL(),dbname=dbname,username=username,password=password,host=host)
```

```{r Results.to.SQL,echo=FALSE,warning=FALSE,error=FALSE,message=FALSE,results=FALSE}

```

```{r, constants,echo=FALSE,results=FALSE,warning=FALSE,error=FALSE,message=FALSE}
SCC <- 24
```

Climate change is one of the most pressing issues facing mankind.
In order to fight climate change, we have to find the most cost-effective ways to curb the emission of GreenHouse Gases (GHGs).
A lot of different actions can be taken to combat climate change.
Among them are interventions on: 

* Buildings (weatherization)
* Transportation (taxes, norms)
* Land use ([forest](ForestPES.html) and [grassland subsidies](GrasslandPES.html))
* Power generation (taxes and markets for polluting rights)

Allocating the amount of public and private money to the most effective uses requires an estimate of the cost-effectiveness of the various options: how much do they cost per ton of C02 saved from the atmosphere?
This monetary estimate is called the abatement cost of the option.

Below is a graph comparing the estimated abatement costs of the various options with the [social cost of carbon](SCC.html):
```{r CE.CC,echo=FALSE,warning=FALSE,error=FALSE,message=FALSE,results=FALSE,fig.align='center',fig.cap='Abatement Cost of Various Options to Fight Climate Change  (in Euros/tonC02)'}
Meta.CC <- dbReadTable(SKY.db,'Meta_CC')
ggplot(Meta.CC, aes(x=program, y=coef)) + 
  geom_pointrange(aes(ymin=coef-1.96*se.coef, ymax=coef+1.96*se.coef), width=.1) +
  geom_hline(aes(yintercept=SCC),color='red',linetype='dashed')+
  annotate('text',x=0.75,y=30,label='Social Cost of Carbon',color='red')+
  expand_limits(y=c(0))+
  ylab("Estimated Abatement Cost (in Euros/tonC02)")+
  xlab("")+
  theme_bw()
```

It appears that the French Grassland Conservation Program is currently not a cost-effective way to fight climate change: its abatement cost is largely above the social cost of carbon.