---
title: "Grassland Conservation and Climate Change"
---

```{r libraries,echo=FALSE,results=FALSE,warning=FALSE,error=FALSE,message=FALSE}
library(gsheet)
library(metafor)
```

Under certain conditions, grassland can avoid the emission of GHGs.
A case in point is when grassland is extensively managed and replaces cropland.
In that case, [it has been computed that](https://hal.archives-ouvertes.fr/hal-01152980/document) 10.55tC02 are stored per hectare of extensively managed grassland saved from conversion to cropland. 
As a result, Grassland Conservation Programs, that pay farmers for maintaining grassland cover, might be an effective way to combat climate change, if they succeed in triggering an increase in grassland cover for a reasonable amount of money.

In this section we report on the results of several studies that have estimated the cost-effectiveness of the French implementation of the EU Grassland Conservation Program, one of the largest Grassland Conservation Programs in the world.
The studies are:

* Chabé-Ferret and Subervie (2009) using observational methods
* Chabé-Ferret and Voia (2019) using the 2000 reform of the program as a natural experiment
* Gallic and Marcus (2018) using the 2015 reform of the program as a natural experiment

```{r download.data,echo=FALSE,results=FALSE,warning=FALSE,error=FALSE,message=FALSE}
data.GCP <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1RKZZuAAnGMUKgoFTQskVSCoMsOxzpcPp-RCZbHgvsK8/edit?usp=sharing")
data.GCP
colnames(data.GCP) <- c("paper.id","authors","year","cost","se.cost")
```

```{r FE.meta,echo=FALSE,warning=FALSE,error=FALSE,message=FALSE,results=FALSE}
data.GCP$var.cost <- data.GCP$se.cost^2
meta.GCP.FE <- rma(yi = data.GCP$cost,vi=data.GCP$var.cost,method="FE")
summary(meta.GCP.FE)
```

The overall estimate is of `r round(coef(meta.GCP.FE),0)` $\pm$ `r round(1.96*summary(meta.GCP.FE)$se,0)` Euros per ton of C02. 
This is still much higher than the social cost of C02 of 24 Euros per ton of C02 as estimated by the US EPA.
It is close to the most recent estimates of the social cost of carbon stemming from the Quinet report.

### Meta-Analysis of Individual Studies

```{r FE.forest.plot,echo=FALSE,warning=FALSE,error=FALSE,message=FALSE,results=FALSE}
forest(meta.GCP.FE,slab = data.GCP$authors,xlab='Estimated Cost-Effectiveness of Grassland Conservation Programs (in Euros/tonC02)')
```

### Analysis of Publication Bias

There is no evidence of publication bias in this dataset.

```{r FE.funnel.plot,echo=FALSE,warning=FALSE,error=FALSE,message=FALSE,results=FALSE}
funnel(meta.GCP.FE)
```